---
title: "Covid-19 Variant Data"
author: "Sam Altshuler  \nPID: A59010373"
date: "7/14/2022"
output: pdf_document
---

## Load in correct libraries
```{r, message=FALSE}
library(dplyr)
library(lubridate)
library(ggplot2)
```

## Load in the Covid-19 Variant data
Read in the data from a csv. Data downloaded from the California Health and Human Services website.

```{r}
total_data <- read.csv("covid19_variants.csv")
# Preliminary look at the data
head(total_data)
```

## Start to graph the variants over time 
Next we can preliminarily graph the data to see what it looks like.

```{r}
ggplot(total_data, aes(x = date, y = percentage, group = variant_name, color = variant_name))+
  geom_line()
```
Now this looks similar to example graph, but the next step is to remove all rows that have "Total" variants as this row will always be 100%. We also need to remove the "other" rows.  

```{r}
# Filter out all rows that have a variant name of "Total"
no_total <- total_data %>% filter(variant_name != "Total" )

# Filter out all rows with a variant name of "Other"
variants <- no_total %>% filter(variant_name != "Other")

# Graph to ensure that total and other are no longer represented on the graph
ggplot(variants, aes(x = date, y = percentage, group = variant_name, color = variant_name))+
  geom_line()
```
Next we want to change the x axis to 1 month increments that show the 3 letters of the month with the year. This is done by using the lubridate package.
```{r}
# Convert each date in the date column into the lubridate date class.
variants$date <- as_date(variants$date)

#graph again, with the X axis broken into 1 month increments and displaying the month and year
ggplot(variants, aes(x = date, y = percentage, group = variant_name, color = variant_name))+
  geom_line()+
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y")
```
Now that our graph now has the correct axes, let's make it more readable by angling the x axis text so it doesn't overlap. We will also get rid of the X axis title. Lastly, we will add a title to the legend.
```{r}
ggplot(variants, aes(x = date, y = percentage, group = variant_name, color = variant_name))+
  geom_line()+
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y")+ 
  labs(y ="Percentage of sequenced specimens", x = "", color = "Variant" )+
  # Angle the x axis text so that they don't overlap
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
  
```
Now that the labels and axes are nice and readable, let's make the lines a bit thicker to be easier to see and add a title to the graph.
```{r}
ggplot(variants, aes(x = date, y = percentage, group = variant_name, color = variant_name))+
  geom_line(size = 0.75)+ #change the width of the lines
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y")+ 
  labs(y ="Percentage of sequenced specimens", x = "", color = "Variant", 
       title = "COVID-19 Variants in California")+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
  
```
Lastly, we need to add a caption to state where the data is from. This will be the final product!
```{r}
ggplot(variants, aes(x = date, y = percentage, group = variant_name, color = variant_name))+
  geom_line(size = 0.75)+ #change the width of the slide
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y")+ 
  labs(y ="Percentage of sequenced specimens", x = "", color = "Variant", 
       title = "COVID-19 Variants in California", 
       # Adds in a caption on the bottom right of the graph to show the data source
       caption = "Data Source: <https://data.chhs.ca.gov/>")+  
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
  
```

